---
title: "PES University, Bangalore"
subtitle: "Established under Karnataka Act No. 16 of 2013"
author:
- 'SRUSHTI S N, Dept. of CSE - PES2UG20CS352'
output: pdf_document
urlcolor: blue
editor_options: 
  markdown: 
    wrap: 72
---

### 

### Loading the dataset

```{r}
library(ggpubr)
library(dplyr)
df <- read.csv('road_accidents_india_2016.csv', row.names=1)
```

### Problem 1 (2 points)

Find the total number of accidents in each state for the year 2016 and
display your results. Make sure to display all rows while printing the
dataframe. Print only the necessary columns. (Hint: use the grep command
to help filter out column names).

```{r}
acc_cols <- grep("Total.Accidents$", colnames(df), ignore.case=T, value=TRUE)

total_accidents <- data.frame(state..ut=df$State..UT, total_acc=rowSums(df[ , c(acc_cols)], na.rm=TRUE))

print.data.frame(total_accidents)


```

### 

Problem 2 (2 points)

Find the (fatality rate =total number of deaths/total number of
accidents) in each state. Find out if there is a significant linear
correlation at a significance of α = 0.05 between the fatality rate of a
state and the mist/foggy rate (fraction of total accidents that happen
in mist/foggy conditions). Plot the fatality rate against the mist/foggy
rate. (Hint: use the ggscatter library to plot a scatterplot with the
confidence interval of the correlation coefficient).

```{r}
death_cols <- grep("Persons.Killed$", colnames(df), ignore.case=T, value=TRUE)
total_accidents$total_deaths <- rowSums(df[ , c(death_cols)])

total_accidents$fatality_rate <- total_accidents$total_deaths             /total_accidents$total_acc

total_accidents$mist_rate <- df$Mist..Foggy...Total.Accidents/total_accidents$total_acc

print.data.frame(total_accidents)

```

Plotting the data

```{r}
plot(x=total_accidents$fatality_rate, y=total_accidents$mist_rate,
col='#00FF00')
```

Finding Pearson's correlation coefficient

```{r}
coeff <- cor(total_accidents$fatality_rate,
total_accidents$mist_rate, method='pearson')
coeff 

corr_test = cor.test(total_accidents$fatality_rate, total_accidents$mist_rate, method='pearson')
print(corr_test)


```

```{r}
ggscatter(total_accidents, x='fatality_rate', y='mist_rate', add='reg.line', conf.int=TRUE, cor.coef=TRUE,col='#00FF00', cor.method = 'pearson')

```

### 

Problem 3 (3 points)

Rank the states based on total accidents and total fatalities (give a
rank of 1 to the state that has the highest value of a property). You
are free to use any tie-breaking method for assigning ranks. Find the
Spearman-Rank correlation coefficient between the two rank columns and
determine if there is any statistical significance at a significance
level of α = 0.05. Also test the hypothesis that the correlation
coefficient is at least 0.2.

```{r}

total_accidents$acc_ranks <- rank(desc(total_accidents$total_acc),
ties.method='random')
total_accidents$death_ranks <- rank(desc(total_accidents$total_deaths),
ties.method='random')


rs <- cor(total_accidents$acc_ranks, total_accidents$death_ranks, method='spearman')
rs
#rs- speraman cofficient

#printing
print(cor.test(total_accidents$acc_ranks, total_accidents$death_ranks, method='spearman'))

ggscatter(total_accidents, x='acc_ranks', y='death_ranks',
add='reg.line',col='#800080', conf.int=TRUE,
cor.coef=TRUE, cor.method = 'spearman')
```

t test

```{r}
degrees <- nrow(total_accidents) - 2
t_stat <- (rs - 0.2)/sqrt((1 - rs*rs)/(nrow(total_accidents) - 2))
t_stat

2 * pt(q=t_stat, df=degrees, lower.tail=FALSE)

```

### 

Problem 4 (1.5 points)

Convert the column Hail.Sleet\...Total.Accidents to a binary column as
follows. If a hail/sleet accident has occurred in a state, give that
state a value of 1. Otherwise, give it a value of 0. Once converted,
find out if there is a significant correlation between the
hail_accident_occcur binary column created and the number of rainy total
accidents for every state. Calculate the point bi-serial correlation
coefficient between the two columns. (Hint: it is equivalent to
calculating the Pearson correlation between a continuous and a
dichotomous variable. You could also use the ltm package's biserial.cor
function).

```{r}

total_accidents$hail_binary <- ifelse(df$Hail.Sleet...Total.Accidents > 0, 1, 0)
total_accidents$rain_acc <- df$Rainy...Total.Accidents
print.data.frame(total_accidents[, c('state..ut', 'hail_binary', 'rain_acc')])

cor.test(total_accidents$rain_acc, total_accidents$hail_binary, method='pearson')
```

From the above values we can see that there isn't any correlation

### 

Problem 5 (1.5 points)

Similar to in Problem 4, create a binary column to represent whether a
dust storm accident has occurred in a state (1 = occurred, 0 = not
occurred). Convert the two columns into a contingency table. Calculate
the phi coefficient of the two tables. (Hint: use the psych package).

```{r}
total_accidents$dust_binary <- ifelse(df$Dust.Storm...Total.Accidents > 0, 1, 0)

contingency_table <- table(total_accidents[, c('dust_binary', 'hail_binary')])

contingency_table

```

```{r}
library(psych)

phi(contingency_table)
```
