---
title: "PES University, Bangalore"
subtitle: "Established under Karnataka Act No. 16 of 2013"
author:
- 'SRUSHTI S N, Dept. of CSE - PES2UG20CS352'
output: pdf_document
urlcolor: blue
editor_options: 
  markdown: 
    wrap: 72
---

## 

# **Worksheet 2a - Simple Linear Regression**

### Loading dataset and inserting libraries

```{r}
library(tidyverse)
library(ggplot2)
library()
#library(dplyr)

df<- read.csv("dragon_neurons.csv")
```

### Problem 1

```{r}
#Problem 1
#estimating using ordinary least square
#finding slope and intercept
#explanatory var: independent var
slope <- cor(df$'axon_diameter', df$'conduction_velocity') * (sd(df$'conduction_velocity') /
                                                                sd(df$'axon_diameter'))
inter <- mean(df$'conduction_velocity') - (slope * mean(df$'axon_diameter'))
df %>%
  ggplot(aes(x = sqrt(df$axon_diameter), y = sqrt(df$conduction_velocity))) +geom_point(colour
                                          = "green")+geom_smooth(method = "lm", fill = NA)



```

Thus linear model is suitable for representing above relationship

### Problem 2

```{r}
#Problem 2
x<-sqrt(df$axon_diameter)
y1<-sqrt(df$conduction_velocity)
df$conduction_velocity2 <- df$conduction_velocity * df$conduction_velocity
y2<-df$conduction_velocity2

model1<-lm(y1~x)
resi<-resid(model1)
qqnorm(resi)
qqline(resi)


model2<-lm(y2~x)
residu<-resid(model2)
qqnorm(residu)
qqline(residu)

```

No, the residuals don't look like white noise. Hence, we transform y by
taking square of it

### Problem 3

```{r}
#problem3
dataa<-df[c("axon_diameter","conduction_velocity2")]
dataa.center  = colMeans(dataa)
dataa.cov     = cov(dataa)

# Ellipse radius from Chi-Sqaure distrubiton
rad  = qchisq(p = 0.95 , df = 2)

# Square root of Chi-Square value
rad  = sqrt(rad)

# Finding ellipse coordiantes
ellipse <- car::ellipse(center = dataa.center , shape = dataa.cov , radius = rad,segments =
                          150 , draw = FALSE)
library(ggplot2)

# Ellipse coordinates names should be same with air data set
ellipse <- as.data.frame(ellipse)
colnames(ellipse) <- colnames(dataa)

# Create scatter Plot
figure <- ggplot(dataa , aes(x = axon_diameter , y = conduction_velocity2)) +
       geom_point(size = 2) +
       geom_polygon(data = ellipse , fill = "orange" , color = "orange" , alpha = 0.5)+
       geom_point(aes(dataa.center[1] , dataa.center[2]) , size = 5 , color = "blue") +
       geom_text( aes(label = row.names(dataa)) , hjust = 1 , vjust = -1.5 ,size = 2.5 ) +
       ylab("conduction_velocity2 Values") + xlab("axon_diameter Values")
       
# Run and display plot
figure
```

```{r}
distances <- mahalanobis(x = dataa , center = dataa.center , cov = dataa.cov)
# Cutoff value for ditances from Chi-Sqaure Dist. 
# with p = 0.95 df = 2 which in ncol(air)
cutoff <- qchisq(p = 0.95 , df = 2)

## Display observation whose distance greater than cutoff value
dataa[distances > cutoff ,]
```

From the plot above and the data, we can conclude that there are no
outliers.

### Problem 4

```{r}
#problem4
summary(model1)
```

Multiple R-squared: 0.8152, Adjusted R-squared: 0.8123. functional form
linear model

```{r}
summary(model2)
```

Multiple R-squared: 0.7823, Adjusted R-squared: 0.779. Higher value of R
2 implies better fit

### Problem 5

```{r}
summary(model2)
```

As we can see that p value is lesser, it is a statically significant
relationship which implies positive relationship between the columns
"axon_diameters" and "conduction_velocity".
