---
title: "PES University, Bangalore"
subtitle: "Established under Karnataka Act No. 16 of 2013"
author:
- 'SRUSHTI S N, Dept. of CSE - PES2UG20CS352'
output: pdf_document
urlcolor: blue
editor_options: 
  markdown: 
    wrap: 72
---

## Time Series Data and Basic Forecasting Techniques

### Loading Dataset and Packages

```{r}
library(tidyverse)
library(fpp2)
df <- read.csv('sales.csv')
head(df)
```

Pick out the Sales column in the data.frame. Most of our time-series
analysis will be done on this column.

```{r}
sales <- df$Sales
head(sales)
```

The ts function is used to create the ts object in R. Frequency is 52 as
it is weekly data. The start is specified like start= c(y, m, d) as we
are dealing with weekly data.

```{r}
sales_ts <- ts(sales, frequency = 52, start=c(2010, 2, 5))
sales_ts

```

Visualize the Time-Series of Sales column

```{r}
plot.ts(sales_ts)

```

#### Problem 1

Decompose the Sales column into trend, seasonal and random components.
Plot these components as well

```{r}
decomp<-decompose(sales_ts)  
plot(decomp)
```

#### Problem 2

• Perform forecasts using Single, Double and Triple Exponential
Smoothing.

• Plot the forecasts of all three forecasts (different colours) against
the true values. (Hint: use lines)
• Use only one function needed for all 3 forecasts, only changing
parameters to get each of the 3 models (Hint: Think about alternate
names)

```{r}
#Creating training data and testing data
sales_train<-window(sales_ts,end=c(2020,40))
sales_test<-window(sales_ts,start=c(2011,2))

```

```{r}
#Simple Exponential Smoothing
#alpha=0.2 is where the model works best
simple<- forecast(ses(sales_train,alpha=0.2,h=100))
autoplot(simple)
```

```{r}
#Double Exponential Smoothing
#Use Holt's method
holt_double<-forecast(holt(sales_train, h=100))
autoplot(holt_double)
```

```{r}
#Triple Exponential Smoothing
hw_triple<-HoltWinters(sales_ts,alpha=0.2, beta=0.5, gamma=0.8, seasonal = "additive")
plot(hw_triple)
triple<-forecast(hw_triple)
plot(triple)
```

#### 
Problem 3

• Forecast Sales values by Regression using all other columns. Print
summary of regression model.
• Plot the predicted values against original as well. (Hint: Regression
model predictions will not be a Time Series, so need to get both to
common index/x-axis)

```{r}
model<- lm(sales ~ (Holiday_Flag + Unemployment + Laptop_Demand + Temperature + Fuel_Price + CPI),data = df)
pred = predict(model)
summary(model)
plot(pred)
plot(predict(pred),df$Sales,ylab="Observed",xlab="Predicted")
abline(0,1)
```

#### Problem 4

Plot the Laptop_Demand column as a time series. Identify the forecasting
required for this type of Time-series,and forecast the values for all
143 weeks (Hint: Look at functions in the forecast package)

```{r}
#Holt winter's method
demand<- ts(df$Laptop_Demand, frequency = 23, start=c(2010, 2, 5))
#plot.ts(demand)
demand_holt <- HoltWinters(demand)
plot(demand, ylab="laptop_sales", xlim=c(2010,2013))
lines(demand_holt$fitted[,1])
```

We can see from above graph how the actual and smoothed model DO NOT fit
together well. Thus using regression is not a good idea.

#### Problem 5

Evaluate the accuracy of all 3 Exponential Smoothing models (from
Problem 2) and Regression models using the MAPE and RMSE metrics.
Comment on which is the best Exponential Smoothing method, and if
Regression is better than Exponential Smoothing? Provide a reasoning for
why the best model is better suited for the Sales data (Bonus Point:
reasoning for why the 2 other models perform similarly)

```{r}
#finding accuracy
cat("Simple")
simple_acc<-accuracy(simple)
simple_acc

cat("Double")
double_acc<-accuracy(holt_double)
double_acc


cat("Triple")
triple_acc<-accuracy(forecast(triple))
triple_acc

cat("Regression model")
model_acc<-accuracy(model)
model
```

We see from the above results that the errors are the least in case of
Triple Exponential Smoothing.Hence, it is the best suitable method and
no, regression is not better than Exponential Smoothing.

Triple Exponential Smoothing involves 3 equations and thus can handle
seasonality and trend which contribute to the reasoning as why Triple
exponential smoothing is the best method.

Bonus: Simple exponential smoothing behaves similar to Triple
Exponential Smoothing but it cannot handle seasonality or trend.
And , Double exponential smoothing behaves similar to Triple Exponential
Smoothing, including handling trend but not seasonality.
